\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-06-27 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\usepackage{lipsum}
\usepackage{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}

\begin{document}

We would like to thank the Editor in Chief and the two reviewers for their helpful comments and reviews. We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we feel that they are not a reasonable request to include in our paper.  We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses below.\\\

\textbf{Response to Associate Editor:}

For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.  However, I took a more negative view towards several aspects of the Methods, Results, and Discussion.  Three examples:

1) I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don’t provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.

2) Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken? Second, why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class.  Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.  Fourth, what was the purpose of the Phenotyping 2
experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.

3) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.  It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus.

Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.

Yours sincerely,
Andrew Schnabel
Associate Editor

Reviewer #1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1. The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis. The authors go on to sequence regions up and downstream of Hopscotch in subset of their initial sample. In analyzing their sequence data, they find no evidence for recent introgression of the maize Hopscotch locus into wild teosintes. They also find evidence for selection acting on tb1 in parviglumis. In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch.

They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate. However, I feel like they need to make some further effort to understand what is happening mechanistically. What is Hopscotch doing in maize vs these sampled teosintes? I feel like the expression analyses and phenotyping experiments they suggest should be part of a later study should be part of this study. At a minimum, qRT-PCR experiments should be performed assessing tb1 expression in their with/without Hopscotch lines. Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results.

Another potential avenue, that the authors also suggest, lies in more directed and exhaustive greenhouse experiments. Variation in red and far-red light has been shown to be important in regulating both tb1 (in Sorghum) and gt1 in maize. Growth chamber experiments (with or without supplemental far red light, with or without Hopscotch), coupled to expression analyses in seedlings, seem like they should be possible.  In the greenhouse experiments that were done, phenotyping was incomplete. Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture and internode length. Was there any evidence for variation in these other traits? In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize.

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown.

Reviewer #2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana. Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.
Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:
Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?
Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…"
Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?
Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?
Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.
Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.
Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".  Sorry, what does this mean? That you have one band (one assumes the short one?). If you are talking presence/absence, how do you know the allele?
Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include?
Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis? Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago?
Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte?
Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb?
Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically? Or you turned on the drip by hand? In which case, it was not automatic…
Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1. F" So, why did you do it? Because of the last sentence in the materials and methods? Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it.
Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols?
Results:
Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total? So we can see if this comes out to a very different ratio than in maize.
Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles? I thought you only had presence or absence; how are alleles differentiated with your PCR?
Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011). Add Warburton, et al., 2011.  Gene flow among different teosinte taxa and into the domesticated maize gene pool.  Genet Resour Crop Evol 58:1243-1261.
Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32.
\end{document}

### 2014-06-28 ResponseToReview.tex

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry} %package that allows changes in margins and header/footers
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\usepackage{lipsum}
\usepackage{color}
\usepackage{multirow}
\renewcommand{\baselinestretch}{1.5}

\begin{document}

\noindent Dear Dr. Andrew Schnabel and anonymous reviewers, \\

We would like to thank the Editor in Chief and the two anonymous reviewers for their helpful comments and suggestions. We have addressed the comments to the text, and while we feel that the suggested new experiments would provide interesting results, we do not think that they are a reasonable request to include in resubmission of our paper.  We have attached a revised version of our manuscript, including the suggested changes, and have detailed out our responses in blue below. We hope that you are willing to consider our revised manuscript for resubmission.\\\

\noindent Sincerely,\\

\res{would this be jeff, matt, me? or everyone?}

\section*{Response to Associate Editor:}

@@

For my own part, I agree with both reviewers that the study would be of interest to readers of AJB.  However, I took a more negative view towards several aspects of the Methods, Results, and Discussion.  Three examples:

1) I thought that the neighbor-joining analysis was of little value, because it was not clear to me just how the result could be expected to provide information about Hopscotch genotype.  Why would you expect individuals with similar Hopscotch genotypes to cluster together?  Also, although you don’t provide any information on the distribution of variable sites among accessions and taxa, your data set of only 48 and 40 segregating sites is almost certainly not sufficiently variable to resolve relationships among the very large number of accessions you were analyzing.  As a result, you produce mostly a huge polytomy, with a small amount of structure that appears to reflect the genetic similarities and differences already evident in Table 1.  The trees themselves are difficult for a reader to interpret, because none of the accession names at the branch tips (other than TIL) are defined in the text.

\res{I vote we take out trees}

2) Phenotyping experiments - These are curious experiments.  First, why did you not generate a population of individuals with known genotypes and then test for the effect of genotype on phenotype?  Why take a random sample and just hope that you will get the necessary number of individuals of each genotype?  \res{we didn't know genotypes ahead of time...they were seeds collected from natural population} How can you randomize the growing conditions with respect to genotype, if you don’t know genotypes ahead of time but instead determine them after all the growing measurements are taken? Second, why are sample sizes not reported?  You provide total samples sizes (eg, 206 or 216 for Phenotyping 1, depending on whether you read the Methods or the Results), but not sample sizes for each genotypic class \res{I will check this. I thought it was listed in the text. maybe older version}.  Third, it was not clear to me what “to detect the observed effect” means in lines 12-13 of p. 9.  Your Phenotyping 1 results show no effect of genotype on tillering.  Thus, it is not clear what effect you are referring to.  Fourth, what was the purpose of the Phenotyping 2 \res{we did phenotyping 2 because we had such an odd distribution of genotypes in Phenotyping 1 that we had no power to make any inferences regarding effect of Hop on TI} experiment?  You already showed no effect of genotype on phenotype using populations where you are most likely to see segregation for the insertion, so it was not clear to me why sampling more broadly, especially from populations {We didn't sample an additional population aside from SLO for phenotyping -- is that what he is inferring?} with no evidence of the Hopscotch insertion, would be expected to improve your understanding of this relationship.

3) The discussion of introgression, genetic drift, and selection in your Discussion section seems to lack coherence.  It sometimes focuses on explaining the unexpectedly high frequency of Hopscotch in present-day populations, whereas other times it seems to be addressing the presence or absence of Hopscotch in teosinte in general, and it also sometimes seems to be addressing selection on the tb1 locus. \res{should we restructure this? I thought it read clearly. Key points were 1) Hop more widely spread than previously thought; 2) evidence of selection on Hop in teo}

### 2014-06-28 ResponseToReview.tex

Because of what I perceive to be substantial problems both with the methods and the presentation and discussion of results, and because Reviewer 1 raises some important issues with the phenotyping experiments, I am recommending that the manuscript be rejected.

Yours sincerely,
Andrew Schnabel
Associate Editor

\section*{Reviewer 1:}

Reviewer #1: In their paper 'Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)', Vann et al survey a large sample of maize and teosinte individuals for the presence of the Hopscotch transposable element upstream of tb1. The authors find that Hopscotch is more widespread, across a large sampled area in Mexico, than previously thought, particularly in parviglumis. The authors go on to sequence regions up and downstream of Hopscotch in subset of their initial sample. In analyzing their sequence data, they find no evidence for recent introgression of the maize Hopscotch locus into wild teosintes. They also find evidence for selection acting on tb1 in parviglumis. In greenhouse experiments, no difference in tiller index or tiller number was observed between teosinte lines with or without Hopscotch. \res{yes sounds right}

They present an interesting result, their conclusions are supported by their data, and the methods they use are appropriate. However, I feel like they need to make some further effort to understand what is happening mechanistically. What is Hopscotch doing in maize vs these sampled teosintes? I feel like the expression analyses and phenotyping experiments they suggest should be part of a later study should be part of this study. At a minimum, qRT-PCR experiments should be performed assessing tb1 expression in their with/without Hopscotch lines. Expression analyses of gt1, te1, and tru1 would be great, since they invoke variation at these loci to explain their results. \res{sure this would all be cool and interesting, but not going to happen for this paper. it is enough work on it's own to be a separate paper}.

Another potential avenue, that the authors also suggest, lies in more directed and exhaustive greenhouse experiments. Variation in red and far-red light has been shown to be important in regulating both tb1 (in Sorghum) and gt1 in maize. Growth chamber experiments (with or without supplemental far red light, with or without Hopscotch), coupled to expression analyses in seedlings, seem like they should be possible.  In the greenhouse experiments that were done, phenotyping was incomplete. Apart from tillering, Hopscotch and natural variation at the tb1 locus appears to affect ear architecture \res{we can't do this unless plants all go to seed} and internode length. Was there any evidence for variation in these other traits? In the longer term, it would be nice to see what happens when you introgress these particular parviglumis loci into maize. \res{no more phenotypine experiments}.

One minor issue: in Fig. 1, it would be nice to see where the Balsas River Basin is in relation to the sampling shown. \res{is this necessary?}

Reviewer #2: Although this paper brings up an interesting question regarding the role of tb1 standing variation in teosinte, it is not able to answer it completely.  It presents results that have been published previously about the presence of hopscotch in teosinte, maize, and mexicana. \res{sure we knew it existed in some individuals, but not to the extent that we found in our study} Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.
This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied. \res{okay. but it is a paper about maize...}
Methods and statistics were appropriately used as far as my expertise allows me to judge.
Specific suggestions follow:

Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples? \res{okay can do this.}

Page 4, Line 9: in the sentence "But while some eﬀort has been invested into", delete the word "But" and start the sentence "While some effort…" \res{doable}

Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte? \res{doable}

Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions? \res{no mexicana phenotyping. seed limited. there were a couple populations of mex that had hop at intermediate frequency.}

Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements. \res{no way we can list all those numbers in the text. varied anywhere from 2 to 30 is my guess. can say that?}

Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements. \res{i think this is obvious in the methods when we discuss primers sets}

Page 6 line 6: "When only one PCR resolved well, we scored one allele for the individual".  Sorry, what does this mean? That you have one band (one assumes the short one?). If you are talking presence/absence, how do you know the allele? \res{PCRs for different alleles. maybe should say that in the text to clarify?}

Page 7 lines 14 - 15 "These analyses only included populations in which eight or more individuals were sampled."  How many populations did that include? \res{can put this in text}

Page 7 lines 17 - 26: how many year's worth of environmental data went into your analysis? Can you hypothesize that tb1 differences can be correlated with current climactic data if the changes and subsequent selections must have happened many generations ago? \res{i have no idea. do either of you know if this is in tanja's paper? or should i look at the website she got data from?}

Page 7 lines 22 - 23: "We used genotyping and covariance data from Pyhajarvi et al. (2013) for BayEnv" what kind of dataset is this? Teosinte? \res{can explain dataset in paper}

Page 8 lines 20 - 21:" We created haplotype blocks using a custom Perl script that grouped SNPs separated by less than 5kb into haplotypes." Does LD data in your parviglumus support 5Kb? \res{hmm. Matt why did you use 5kb when you did this for your study?}

Page 9 lines 2 - 3: "Plants were watered three times a day by hand and with an automatic drip containing 10-20-10 fertilizer." This sounds odd; by hand or automatically? Or you turned on the drip by hand? In which case, it was not automatic… \res{i'll just take out the by hand. it's too confusing to explain what happened}.

Page 9 lines 7 - 8: "Culm diameter is not believed to be correlated with tillering index or variation at tb1. F" So, why did you do it? Because of the last sentence in the materials and methods? Perhaps do not mention here that is not correlated; wait until the end of the materials and methods where you seem to explain it. \res{we wanted an independent trait to compare patterns of TI to}

Page 9 lines 9 - 10: "We genotyped individuals for the Hopscotch insertion following the protocols listed above." The PCR protocols or the sequencing protocols? \res{PCR}

Results:
Page 10 lines 5 - 6: "Within our parviglumis and mexicana samples we found the Hopscotch insertion segregating in 37 and four populations, respectively"; remind us how many parviglumis and Mexicana samples you had total? So we can see if this comes out to a very different ratio than in maize. \res{okay can put in ratios or something}

Page 10 lines 7 - 9: "Using our Hopscotch genotyping, we calculated diﬀerentiation between populations (FST) and subspecies (FCT) for populations in which we sampled eight or more alleles." What do you mean, 8 or more alleles? I thought you only had presence or absence; how are alleles differentiated with your PCR? \res{can clarify. basically only used populations that had aboive a certain sample size. alleles based on size difference in PCR}.

Page 13 lines 25 - 27: "and genetic data support ongoing gene ﬂow between domesticated maize and both mexicana and parviglumis in a number of sympatric populations (Huﬀord et al., 2013; Ellstrand et al., 2007; van Heerwaarden et al., 2011). Add Warburton, et al., 2011.  Gene flow among different teosinte taxa and into the domesticated maize gene pool.  Genet Resour Crop Evol 58:1243-1261. \res{okay good suggestion}.

Page 14 line 15 to page 15 line 14: too long a paragraph. Break into two, perhaps at page 14 line 32. \res{okay}.

### 2014-07-07 manuscript.tex
Edits
Edits to text and response letter

\author[2]{Laura Vann}%author information
\author[2,3]{Thomas Kono}
\author[2,4]{Tanja Pyh\"aj\"arvi}
\author[2,5,7]{Matthew B. Hufford}
\author[2,6,7]{Jeffrey Ross-Ibarra}
\affil[2]{Department of Plant Sciences, University of California Davis}
\affil[3]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
\affil[4]{Department of Biology, University of Oulu}
\affil[5]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
\affil[6]{Center for Population Biology and Genome Center, University of California Davis}
\affil[7]{Author for Correspondence}
\date{}

@@

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 216 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium. After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch}  insertion and tillering index (Fig. \ref{Fig4Boxplots}), tiller number, or culm diameter.

We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.  For the second grow-out we measured tillering index every five days through day 50 for 302 plants. We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.6 frequency of \emph{Hopscotch} homozygotes, and a 0.2 frequency of both heterozygotes and homozygotes for the teosinte allele. Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.

>>>>

\author[1]{Laura Vann}%author information
\author[1,2]{Thomas Kono}
\author[1,3]{Tanja Pyh\"aj\"arvi}
\author[1,4,6]{Matthew B. Hufford}
\author[1,5,6]{Jeffrey Ross-Ibarra}
\affil[1]{Department of Plant Sciences, University of California Davis}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
\affil[3]{Department of Biology, University of Oulu}
\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
\affil[5]{Center for Population Biology and Genome Center, University of California Davis}
\affil[6]{Author for Correspondence}
\date{}

@@

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. We measured tillering index (TI), the ratio of the sum of tiller lengths to plant height, for 206 plants (Phenotyping 1) from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. Our plantings produced 20 homozygotes for the teosinte (no Hopscotch) allele, 104 heterozygotes, and 82 homozygotes for the maize (Hopscotch) allele. We found the \emph{Hopscotch} segregating at a frequency of 0.65 with no significant deviations from expected frequencies under Hardy-Weinberg equilibrium. After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no correlation between genotype at the \emph{Hopscotch}  insertion and tillering index (Fig. \ref{Fig4Boxplots}), tiller number, or culm diameter.

We performed a second grow-out  of \emph{parviglumis} from San Lorenzo (Phenotyping 2) to assess whether lighting conditions or sample size may have affected our ability to detect an effect of \emph{tb1}.  For the second grow-out we measured tillering index every five days through day 50 for 247 plants. We found the \emph{Hopscotch} allele segregating at a frequency of 0.69, with a 0.56 frequency of \emph{Hopscotch} homozygotes (138 individuals), a 0.2 frequency of homozygotes for the teosinte allele (49 individuals) and a 0.24 frequency of heterozygotes (60 individuals). Results were similar to Phenotyping 1, with no significant correlation between \emph{Hopscotch} and any of the three phenotypes measured.
